---
title: "Final Project"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
# Libraries
library(lme4)
library(olsrr)
library(lmtest)
library(ggpubr)
library(ggplot2)
library(survival)
library(ggfortify)
library(survminer)
library(tidyverse)
library(ggfortify)
library(ggcorrplot)
library(tidymodels)
```


# Phase 1 - (40%)
- The company want to create a model that predicts the probability of a New Client to repurchase in the Month 5, based on their transactional behavior during the first month (considering the first month as M0).

```{r}
# Loading dataset with the first month activity for every client
data1 <- readxl::read_excel("./Data/clients_first_month_VF.xlsx", sheet = "db")

# Changing the variables types
data1$id <- as.numeric(data1$id)
data1$age_gen <- as.factor(data$age_gen)
data1$date_order <- as.Date(data$date_order)
data1$MXN <- as.numeric(data$MXN)
data1$products <- as.numeric(data$products)

head(data1)
```

```{r}
# Loading dataset with the first month activity for every client
data2 <- readxl::read_excel("./Data/clients_18months_VF.xlsx", sheet = "db")

# Changing the variables types
data2$id <- as.numeric(data2$id)
data2$month <- as.factor(data2$month)
data2$MXN <- as.numeric(data2$MXN)
  
head(data2)
```


## 1.1) Aditional variables to calculate:
  - Recency: Days since last order considerig the end of the period as 31/03/2021.
  - Frequency: Number of orders in the first month.
  - Order Size: AVG MXN of the orders in their first month
  - Total MXM: Total MXN spend in the first month.
```{r, message=FALSE}
# Making the variables for the first dataset
data1 <- data1 %>% 
  group_by(id) %>% 
  summarise(max_month = max(date_order),
            recency = as.numeric(difftime(as.Date("2021-03-31"), max_month, units = "days")),
            frequency = as.numeric(n()),
            order_size = as.numeric(mean(MXN)),
            total_MXN = as.numeric(sum(MXN))
            )

data1 <- data1 %>% select(-max_month)
data1
```

```{r}
# Making the survive variable
data1_5 <- data2 %>% 
  filter(month == "M5") %>% 
  mutate(status = ifelse (MXN > 0, 1, 0))

# Selecting the columns that we need
data1_5 <- data1_5 %>% select(id, status)
data1_5
```

```{r}
# Merging both dataset
data1 <- merge(data1, data1_5, by="id")
data1
```

## 1.2) Develop a data exploration.



## 1.3) Develop a logistic regression in order to solve the problem. (The split of train and test is up to you)
```{r}
# Train and test
set.seed(10)

split_inicial<- initial_split(
                data = data1,
                prop = 0.8,
                strata = status
                )

train <- training(split_inicial)
test <- testing(split_inicial)

# Training the model
logistic <- glm(status ~ ., data=train, family="binomial")
summary(logistic)
```
  
- Interpret the coefficients of your model.

  
- Review the performance of the model on the test data ( Accuracy, Sensitivity, Specificity, Confusion Matrix)
  - Consider as threshold 40%, 50% and 60%,  present the performance for all the thresholds and select the best one ( Justify your decision)


- Once that the threshold is selected, calculate the probability of repurchase for all the dataset (train and test) and based on the probability clasify the new clients:
  - Above or equal to the threshold selected as "High Probability"
  - Less than 15% as "Low Probability"
  - The rest as "Medium Probability"



# Phase 2 - (40%)
- Generate a survival analysis for the new clients.
- Consider as "dead" those who do not make a purchase in the month 18.

- Use the variables, age generation, market, probability label and recency and interpret the coeficients.

- Generate the survival curves based on the following.
  - Graph 1 - Based on probability label

  - Graph 2 - Based on market

  - Interpret both graphs

